---
title: "STAT431_HW4"
author: "Taiga Hasegawa(taigah2)"
date: "2019/3/9"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question1

(a)
$$E[(\pi-E(\pi|y))^2|Y=y]=\int_0^1(\pi-E(\pi|y))^2p(\pi|y)d\pi=\frac{\int_0^1(\pi-E(\pi|y))^2p(\pi)L(\pi;y)d\pi}{\int_0^1p(\pi)L(\pi;y)}$$
(b)
```{r}
like <- function(pi) pi^12 * (1-pi)^(58)

prior <- function(pi) dbeta(pi, 0.5, 0.5)

posterior_mean=(12+0.5)/(70+0.5+0.5)
```

```{r}
numerator=integrate(function(pi) (pi-posterior_mean)^2 * prior(pi) * like(pi), 0, 1)
denominator=integrate(function(pi) prior(pi) * like(pi), 0, 1)
numerator$value/denominator$value
```

(c)
```{r}
((12+0.5)*(58+0.5))/((12+58+0.5+0.5)^2*(12+58+0.5+0.5+1))
```
I got the same answer with my approximation in (b).

##Question2

(a)
```{r}
x=c(136,119,114,113,113,112,112,108,102)
y=c(120,93,87,85,76)
ybar1=mean(x)
s1=sd(x)
ybar2=mean(y)
s2=sd(y)
```
Sample mean for small car is `r ybar1` and for SUVs is `r ybar2`. Smaple standard deviations for small car is `r s1` and for SUVs is `r s2`.

(b)

```{r}
Nsim=100000
n1=9
n2=5
sigma1.2s = 1 / rgamma(Nsim, (n1-1)/2, (n1-1)*s1^2/2)
sigma2.2s = 1 / rgamma(Nsim, (n2-1)/2, (n2-1)*s2^2/2)

mu1s = rnorm(Nsim, ybar1, sqrt(sigma1.2s/n1))
mu2s = rnorm(Nsim, ybar2, sqrt(sigma2.2s/n2))

quantile(mu1s - mu2s, c(0.025, 0.975))
```
Because this credible interval doesn't include 0, the means for small cars and SUVs are different. 

(c)
```{r}
mean(mu1s < mu2s)
```
Posterior probability that the mean for small cars does not exceed the mean for SUVs is 2.379%.

(d)
```{r}
t.test(mu1s,mu2s,alternative = "greater",  var.equal=FALSE)
```

```{r}
t.test(mu1s,mu2s,alternative = "greater",  var.equal=TRUE)
```
In both Welch t-test and usual t-test, the null hypothesis that mean for small cars does not exceed the mean for SUVs was rejected. This result is consistent with the answer in (b).

(e)
```{r}
quantile(sigma1.2s/sigma2.2s, c(0.025, 0.975))
```

Because this credible interval includes 1, we can say the variances of small cars and SUVs appear to be same.