---
title: "STAT431_HW3"
author: "Taiga Hasegawa(taigah2)"
date: "2019/2/27"
output:
  pdf_document:
    keep_tex: yes
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question1
(a)
```{r}
x=c(481,144,93,446,69,170,383,63,79,181)
logx=log(x)
ybar=mean(logx)
s=sd(logx)
n=length(logx)
mu0=0
kappa=1
```

```{r}
sigma.2.prior.mean=1
sigma.2.prior.var=1
alpha=sigma.2.prior.mean^2 / sigma.2.prior.var + 2
beta=(alpha-1) * sigma.2.prior.mean
```
The prior distribution is normal-inverse gamma distribution. The mean of normal distribution is `r mu0` and variance is `r kappa` and $\alpha$ is `r alpha` and $\beta$ is `r beta`

(b)
```{r}
mu1=(kappa*mu0+n*ybar)/(kappa+n)
tau1square=(alpha + n/2) / (beta + 0.5 * (n-1)*s^2 +0.5 * kappa*n*(mu0-ybar)^2/(kappa + n))
mu.posterior.location=mu1
mu.posterior.scale=1/(tau1square*(kappa+n))
mu.posterior.df=2*alpha+n
```

The posterior marginal distribution of $\mu$ is t distirbution. The location is `r mu.posterior.location`, the scale is `r mu.posterior.scale` and the degree of freedom is `r mu.posterior.df`.

(c)
```{r}
mu.posterior.mean=mu.posterior.location
mu.posterior.var=(mu.posterior.df/ (mu.posterior.df - 2)) * mu.posterior.scale
mu.posterior.location+qt(c(0.025, 1-0.025), mu.posterior.df) * sqrt(mu.posterior.scale)
```

(d)
```{r}
alpha.posterior=alpha + n/2
beta.posterior=beta + 0.5 * (n-1)*s^2 +0.5 * kappa*n*(mu0-ybar)^2/(kappa + n)
```

The posterior marginal distribution of $\sigma^2$ is inverse gamma distribution with $\alpha$ `r alpha.posterior` and $\beta$ `r beta.posterior`.

(e)
```{r}
1 / qgamma(c(1-0.025, 1-0.975), alpha.posterior, beta.posterior)
```

##Question2
(a)
```{r}
mu.location=ybar
mu.scale=s^2/n
mu.df=n-1
```

The posterior marginal distribution of $\mu$ is t distribution. The location is `r mu.location` and scale is `r mu.scale` and degree of freedom is `r mu.df`.

(b)
```{r}
mu.posterior.mean=ybar
mu.posterior.var=(n-1) * mu.scale / (n-3)
ybar + qt(c(0.025, 1-0.025), n-1) * sqrt(mu.scale)
```

(c)
```{r}
alpha.posterior=(n-1) / 2
beta.posterior=(n-1)*s^2/2
```

The posterior marginal distribution of $\sigma^2$ is inverse gamma distribution with $\alpha$  `r alpha.posterior` and $\beta$ `r beta.posterior`.

(d)
```{r}
(n-1) * s^2 / qchisq(c(1-0.025, 1-0.975), n-1)
```

