---
title: "STAT431_HW2"
author: "Taiga Hasegawa(taigah2)"
date: "2019/2/13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question1

(a)

The posterior density function for uniform prior is 

$\frac{\Gamma(72)}{\Gamma(7)\Gamma(65)}\pi^{6}(1-\pi)^{64}$.


```{r}
n=70
y=6
dx <- 1/1000  # pi grid spacing
pi.grid <- seq(0, 1, dx)
flat.prior <- dunif(pi.grid,0,1)
post.unscaled <- flat.prior * dbinom(y,n,pi.grid)
posterior <- post.unscaled / sum(post.unscaled * dx)
plot(pi.grid, posterior, type="l",
     xlab=expression(pi), main="Posterior: Flat Prior")
```

The posterior density function for beta prior is 

$\frac{\Gamma(170)}{\Gamma(106)\Gamma(164)}\pi^{105}(1-\pi)^{163}$.

```{r}
beta.prior <- dbeta(pi.grid,100,100)
post.unscaled <- beta.prior * dbinom(y,n,pi.grid)
posterior <- post.unscaled / sum(post.unscaled * dx)
plot(pi.grid, posterior, type="l",
     xlab=expression(pi), main="Posterior: Beta Prior")
```

(b)
```{r}
#posterior mean when uniform prior 
mean1=(6+1)/(70+1+1)
#postrior mean when beta prior
mean2=(6+100)/(70+100+100)
#posterior standard deviation when uniform prior
std1=sqrt((6+1)*(64+1)/(6+1+64+1)^2*(6+1+64+1+1))
#posterior standard deviation when beta prior 
std2=sqrt((6+100)*(64+100)/(6+100+64+100)^2*(6+100+64+100+1))
```

When we used uniform prior, posterior mean was `r mean1` and standard deviation was `r std1`.
When we used beta prior, posterior mean was `r mean2`  and standard deviation was `r std2`.

(c)
```{r}
#95% equal-tailed credible interval of uniform prior
qbeta(c(0.025, 0.975), y+1, n-y+1)
```

```{r}
#95% equal-tailed credible interval of beta prior
qbeta(c(0.025, 0.975), y+100, n-y+100)
```

(d)
```{r}
#posterior probabilities of H0 of uniform prior
1 - pbeta(0.2, y+1, n-y+1)
```
The null hypothesis was rejected. We are in favor of  alternative.

```{r}
#posterior probabilities of H0 of beta prior 
1 - pbeta(0.2, y+100, n-y+100)
```
The null hypothesis was not rejected . 
