---
title: "STAT431_HW5"
author: "Taiga Hasegawa(taigah2)"
date: "2019/4/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Question1

(a) 

$\sigma^2_1=\sigma^2_2=\sigma^2$ In this case, $Y_i\mid Z, \sigma^2 \sim i.i.d.N(0,1+\sigma^2)$ and Y is exchangeable.

(b) 

$Cov(Y_1,Y_2)=E[Y_1Y_2]=E[(Z+\epsilon_1)(Z+\epsilon_2)]=E[Z^2]=Var(Z)=1$

Z is independent each other.

#Question2
(b)

model {

  for (i in 1:length(psihat)) {
  
    psihat[i] ~ dnorm(psi[i],tausq[i])
    
    psi[i] ~ dnorm(psi0,tausq0)
    
    tausq[i] <- 1 / sigma[i]^2
    
  }

  psi0 ~ dnorm(0,0.001)

  tausq0 ~ dgamma(0.001,0.001)

  sigmasq0 <- 1/tausq0
  
}

(c)
```{r}
library(rjags)
d <- read.table("hw5.txt", header=TRUE)
inits=list(list(psi0=0.1,tausq0=0.0001),list(psi0=5,tausq0=1),list(psi0=10,tausq0=0.01))
m=jags.model("prob2.bug",d,inits,n.chains = 3)
```
```{r}
x <- coda.samples(m, c("psi0","sigmasq0"), n.iter=20000)
gelman.plot(x, autoburnin=FALSE, ask=TRUE)
```
```{r}
x <- coda.samples(m, c("psi0","sigmasq0"), n.ite=150000)
gelman.plot(x, autoburnin=FALSE, ask=TRUE)
```

20000 iterations should be burned out because shrink factor of all variables becomes significantly below 1.05 after 20000 iterations.

(d)
```{r}
summary(window(x, 20001, 170000))
plot(window(x, 20001, 170000), trace=FALSE, ask=TRUE)
```

